---
title: "Obtaining TRI Data"
author: "Marc Los Huertos"
date: "8/8/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## TRI Data

## Numerous Ways to Get Data

Suggestions: 

1. Collect Data from TRI Explorer

## Using TRI Explorer

1. Check out the fact sheet: 

<https://iaspub.epa.gov/triexplorer/tri_factsheet_search.searchfactsheet>

We might check out [Aniline in the USA](https://iaspub.epa.gov/triexplorer/chemical.html?pYear=2015&pLoc=000062533&pParent=TRI&pDataSet=TRIQ1)

Or select a [region (5Cs are in Region 9](https://iaspub.epa.gov/triexplorer/region.html?pYear=2015&pLoc=9&pParent=TRI&pDataSet=TRIQ1))


2. Evaluate Release Reports

- [Chemicals](https://iaspub.epa.gov/triexplorer/tri_release.chemical)
- [Facilities including Long and Lat.](https://iaspub.epa.gov/triexplorer/tri_getcounties.getcounties?report=tri_release.facility&scriptname=facility&state=06&c_year=2015&c_industry=ALL&c_chemical=_ALL_&c_chemlist=&c_coreyear=&c_indlist=&c_usrState=&c_fips=00000&c_tabrpt=1&c_zip=&c_chk0=false&c_chk1=false&c_chk2=false&c_chk3=false&c_chk4=true&c_chk5=false&c_chk6=false&c_chk7=true&c_chk8=false&c_chk9=false&c_chk10=)
 
 
## Downloaded Facility Data

```{r}
LA2015 <- read.csv("/home/CAMPUS/mwl04747/github/Environmental_Hazards/Data/fac_release2015_2.csv"); head(LA2015)

with(LA2015, plot(Longitude, Latitude))
```


##

I have downloaded data using, <https://www.epa.gov/toxics-release-inventory-tri-program/tri-basic-data-files-calendar-years-1987-2014> and selecting 2014 data. 

### Importing the data

```{r cars}
TRI14CAsource <- "/home/CAMPUS/mwl04747/github/Environmental_Hazards/Data/TRI_2014_CA.csv"
TRI14CA <- read.csv(TRI14CAsource); head(TRI14CA)
```

## Filter for LA County
```{r filterLA}
library(dplyr)
TRI14LA <-  TRI14CA %>% filter(COUNTY == "LOS ANGELES")
```

## Map Sites

```{r maplocations, echo=TRUE}
plot(TRI14LA$LATITUDE, TRI14LA$LONGITUDE)
```

## Group by Chemical Type 

