---
title: "Tests for Association"
author: "Marc Los Huertos"
date: "10/25/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Test of Association

### What is a test of Assocication?

- Is one variable related to another
- Based on count data (frequency data)

## Define Matrix Example

```{r cars, echo = TRUE}
frequencies = c(450, 6560, 132, 2534)
asthma = matrix(frequencies, nrow=2, byrow=T,
dimnames = list(c("City A", "City B"), c("Asthmatic", "Non-Asthmatic")))
```

## View of the data

```{r data}
asthma

```
## Percentages...signficance is unclear!

If you live in city A, what is the probability you'll be asthmatic?

`r round(100*asthma[1,1]/sum(asthma[1,]), 1)`%
or not...
`r round(100*asthma[1,2]/sum(asthma[1,]), 1)`%

If you live in city B, what is the probability you'll be asthmatic?

`r round(100*asthma[2,1]/sum(asthma[2,]), 1)`%
 or not...
`r round(100*asthma[2,2]/sum(asthma[2,]), 1)`%



## Stastistical Analysis

```{r fisher}
fisher.test(asthma)
```


## What do the numbers mean?

### `r round(100*asthma[1,1]/sum(asthma), 1)`%  City A Asthmatic

### `r round(100*asthma[2,1]/sum(asthma), 1)`%  City B Asthmatic

###`r round(100*asthma[1,2]/sum(asthma), 1)`%  City A Non-Asthmatic

### `r round(100*asthma[2,2]/sum(asthma), 1)`%  City B Non-Asthmatic

## Define Matrix Example

```{r Smoking, echo = TRUE}
frequencies = c(20, 20, 20, 20)
smoking = matrix(frequencies, nrow=2, byrow=T,
dimnames = list(c("Cancer", "No Cancer"), c("Smoking", "Non-Smoking")))
smoking
```

## Smoking -- Fisher Test

```{r, echo = TRUE}
fisher.test(smoking)
```
