---
title: "Poisson Distribution"
author: "Marc Los Huertos"
date: "11/1/2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

## What are tests of association?

## Chi square tests

The Chi-squared test is used to evaluate categorical count data -- often these are call contingency tests.

### Are Birthdays Randomly Distributed?

Below are the data we collected:
```{r bdays}
Ja = 3; Fe = 2; Mr = 0; Ap = 1
My = 0; Jn = 0; Jl = 0; Ag = 2
Sp = 3; Oc = 1; Nv = 4; De = 3

Bdays = c(Ja, Fe, Mr, Ap, My, Jn, Jl, Ag, Sp, Oc, Nv, De); Bdays
```
## Histogram 

```{r}
hist(Bdays, xlim=c(0,5))
```

## Histogram
Okay, now we'll make a histogram.

```{r, echo=T}
hist(Bdays, las=1)

```
NOTE: the 'las=1' rotates the labels on the y-axis by 90 degrees, to make it easier to read.

## Expectations

What might be the expectation if the dates were randomly distributed? Each month would have an equal probabliity of 1/12.

```{r}
Bdays_expected = rep(1/12, 12); Bdays_expected
```

What is the sum of these expected probabilities?  Hopefully it is 1! Let's check.

```{r}
sum(Bdays_expected)
```

## Frequencies

Let's convert the frequency to probabilities

```{r}
Bdays_sum = sum(Bdays)
Bdays_prob = Bdays/Bdays_sum; Bdays_prob
```

## Let's compare the frequency and the expected:

```{r}

plot(Bdays_expected, Bdays_prob)
abline(0, 1, col="blue", lwd=2)
```

This is a lame graph -- completely uniformative.


## Chi Square

```{r, echo=FALSE}
chisq.test(Bdays_prob, p= Bdays_expected, rescale.p = TRUE)
```

## hist
Let's try the Poisson Distribution, since this is count data!
```{r}
hist(Bdays, las=1, prob=T, xlim=c(0,7))
```

## Let's calculate the mean:

```{r, label = icky}
xbar = mean(Bdays)
sd = sd(Bdays)
```

## Hist
```{r}
x=seq(0,12,1); x
fy = dpois(x, xbar); fy
hist(Bdays, prob=T, xlim=c(0,8), las=1)
lines(x, fy, col="red")
```

## Expected 
```{r}
expected = c(fy*12); expected
expected
observed = c(4,2,2,3,1)
observed[6:13] = 0; observed
```

```{r}
plot(expected, observed, las=1)
abline(0,1)

```
We could try to run a bestfit line and do a regression, such as:

```{r, label=test}
summary(lm(observed~expected))
plot(expected, observed, las=1)
abline(0,1)
abline(coef(lm(observed~expected)))
```

## New Slide

Lots of problems with this approach!

- contegorical versus continuous
- observed values are truncated.

## aaaa
```{r}
observed[6:13] = 0
plot(0:13, pchisq(0:13, 4), ty="b")
lines(0:13, dchisq(0:13, 4), ty="l", col="blue")
chi2 = sum((observed - expected)^2/expected); #chi2

abline(v=chi2, col="red")
pchisq(chi2, 4, lower.tail=F)
Critical = qchisq(.05, 4, lower.tail=F); Critical
abline(v=Critical, col="purple", lwd=2)
```

## sss
```{r}


chisq.test(observed, p=expected, rescale.p = TRUE)
```

why is this a one-tailed test?

## More Poisson

Toms River -- not done yet!

```{r, label=test2}
observed = 230; expected = 215.5; ratio = observed/expected; ratio
observed*(1-(1/(9*observed))-(1.96/(3*sqrt(observed))))^3/expected
(observed+1)*(1-(1/(9*(observed+1)))+(1.96/(3*sqrt(observed+1))))^3/expected
```
